---
title: "HW3"
output: html_document
---

# #1. Doubly Robust Estimator Simulation 

```{r}
library(readr)
library(dplyr)
library(tableone)
library(geepack)
```


```{r}
nhefs <- read_csv("nhefs.csv")

# 82년도에 관측치가 없어 절단된 데이터 : censoring 
table(nhefs$qsmk, is.na(nhefs["wt82"]))
nhefs$cens <- ifelse(is.na(nhefs["wt82"]), 1, 0)
```

## 1. Doubly Robust Estimator 

### (1) Treatment TRUE / Outcome TRUE 

```{r}
## STEP1. IPW for treatment A 
# Estimation of P(A=1|L) (=IP weights) : via logistic regression 

# A : qsmk (treatment)
fit <- glm(qsmk ~ sex + race + age + I(age^2) +
           as.factor(education) + smokeintensity +
           I(smokeintensity^2) + smokeyrs + I(smokeyrs^2) +
           as.factor(exercise) + as.factor(active) + wt71 + I(wt71^2),
           family = binomial(), data = nhefs)
# fit  # coef 
```

```{r}
## STEP2. Fit Outcome Regression E[Y|A=a, L=l, R]
# qsmk = 0 이면 1-p / qsmk=1 이면 p 
p.qsmk.obs <- ifelse(nhefs$qsmk == 0, 1-predict(fit, type = "response"),
                     predict(fit, type = "response"))

nhefs$w <- 1/p.qsmk.obs  # 1629 개의 weight : 1/{P(Ai=1|Li)}, 1/{P(Ai=0|Li)}
summary(nhefs$w)  # 1.06 ~ 16.01 
```

```{r}
# R = W^A if A = 1 ; R = -W^A if A = 0 
nhefs$r <- ifelse(nhefs$qsmk == 1, 1/p.qsmk.obs, -(1/p.qsmk.obs)) 
summary(nhefs$r)  # -2.92 ~ 16.01
```

```{r}
# GEE, Y ~ A+R, weight <- IPW weight 
msm.w <- geeglm(wt82_71 ~ qsmk + r, data=nhefs, id=seqn, weights = w,  
                corstr="independence")
summary(msm.w)
```

```{r}
## STEP3. Use the predicted values from the outcome model to obtain Standardized mean and outcomes 
nhefs$predicted.meanY <- predict(msm.w, nhefs)  # y hat 
summary(nhefs$predicted.meanY[nhefs$cens==0]) # Min 1.07, Max 10.35 / Mean 2.58
```

```{r}
# create a dataset with 3 copies of each subject
nhefs$interv <- -1     # 1st copy: equal to original one

interv0 <- nhefs       # 2nd copy: treatment set to 0, outcome to missing
interv0$interv <- 0    # treatment A = 0 으로 고정 
interv0$qsmk <- 0
interv0$predicted.meanY <- NA

interv1 <- nhefs       # 3rd copy: treatment set to 1, outcome to missing
interv1$interv <- 1    # treatment A = 1 으로 고정 
interv1$qsmk <- 1
interv1$predicted.meanY <- NA

onesample <- rbind(nhefs, interv0, interv1)  # combining datasets : 1629*3 = 4887 
```

```{r}
std <- glm(predicted.meanY ~ qsmk + sex + race + age + I(age*age)
           + as.factor(education) + smokeintensity
           + I(smokeintensity*smokeintensity) + smokeyrs
           + I(smokeyrs*smokeyrs) + as.factor(exercise)
           + as.factor(active) + wt71 + I(wt71*wt71) + I(qsmk*smokeintensity),
           data=onesample) # 3개의 dataset 으로 modeling 
summary(std) # coef 
```

```{r}
onesample$predicted_meanY <- predict(std, onesample)
mean(onesample[which(onesample$interv==-1),]$predicted_meanY) 
```

```{r}
EY.a0.c0<-mean(onesample[which(onesample$interv==0),]$predicted_meanY)
EY.a1.c0<-mean(onesample[which(onesample$interv==1),]$predicted_meanY)
EY.a1.c0 - EY.a0.c0
```


### (2) Treatment FALSE / Outcome TRUE 

```{r}
nhefs <- read_csv("nhefs.csv")

# 82년도에 관측치가 없어 절단된 데이터 : censoring 
nhefs$cens <- ifelse(is.na(nhefs["wt82"]), 1, 0)
```

```{r}
## STEP1. IPW for treatment A 
# Estimation of P(A=1|L) (=IP weights) : via logistic regression 

# A : qsmk (treatment)
fit <- glm(qsmk ~ sex + race + age + I(age^2) +
           as.factor(education) + smokeintensity +
           I(smokeintensity^2) + smokeyrs + I(smokeyrs^2), 
           family = binomial(), data = nhefs)

## STEP2. Fit Outcome Regression E[Y|A=a, L=l, R]
p.qsmk.obs <- ifelse(nhefs$qsmk == 0, 1-predict(fit, type = "response"),
                     predict(fit, type = "response"))

nhefs$w <- 1/p.qsmk.obs
summary(nhefs$w)  # Min 1.06, Max 17.52 / Mean 2.00 
```

```{r}
# R = W^A if A = 1 ; R = -W^A if A = 0 
nhefs$r <- ifelse(nhefs$qsmk == 1, 1/p.qsmk.obs, -(1/p.qsmk.obs)) 
summary(nhefs$r)  # -2.94 ~ 17.52
```

```{r}
# GEE, Y ~ A+R, weight <- IPW weight 
msm.w <- geeglm(wt82_71 ~ qsmk + r, data=nhefs, id=seqn, weights = w,  
                corstr="independence")
summary(msm.w)
```

```{r}
## STEP3. Use the predicted values from the outcome model to obtain Standardized mean and outcomes 
nhefs$predicted.meanY <- predict(msm.w, nhefs)  # y hat 
summary(nhefs$predicted.meanY[nhefs$cens==0])
```

```{r}
# create a dataset with 3 copies of each subject
nhefs$interv <- -1     # 1st copy: equal to original one

interv0 <- nhefs       # 2nd copy: treatment set to 0, outcome to missing
interv0$interv <- 0    # treatment A = 0 으로 고정 
interv0$qsmk <- 0
interv0$predicted.meanY <- NA

interv1 <- nhefs       # 3rd copy: treatment set to 1, outcome to missing
interv1$interv <- 1    # treatment A = 1 으로 고정 
interv1$qsmk <- 1
interv1$predicted.meanY <- NA

onesample <- rbind(nhefs, interv0, interv1)  # combining datasets : 1629*3 = 4887 
```

```{r}
std <- glm(predicted.meanY ~ qsmk + sex + race + age + I(age*age)
           + as.factor(education) + smokeintensity
           + I(smokeintensity*smokeintensity) + smokeyrs
           + I(smokeyrs*smokeyrs) + as.factor(exercise)
           + as.factor(active) + wt71 + I(wt71*wt71) + I(qsmk*smokeintensity),
           data=onesample) # 3개의 dataset 으로 modeling 

onesample$predicted_meanY <- predict(std, onesample)
mean(onesample[which(onesample$interv==-1),]$predicted_meanY) 

EY.a0.c0<-mean(onesample[which(onesample$interv==0),]$predicted_meanY)
EY.a1.c0<-mean(onesample[which(onesample$interv==1),]$predicted_meanY)
EY.a1.c0 - EY.a0.c0
```



### (3) Treatment TRUE / Outcome FALSE 

```{r}
nhefs <- read_csv("nhefs.csv")

# 82년도에 관측치가 없어 절단된 데이터 : censoring 
nhefs$cens <- ifelse(is.na(nhefs["wt82"]), 1, 0)
```

```{r}
## STEP1. IPW for treatment A 
# Estimation of P(A=1|L) (=IP weights) : via logistic regression 

# A : qsmk (treatment)
fit <- glm(qsmk ~ sex + race + age + I(age^2) +
           as.factor(education) + smokeintensity +
           I(smokeintensity^2) + smokeyrs + I(smokeyrs^2) +
           as.factor(exercise) + as.factor(active) + wt71 + I(wt71^2),
           family = binomial(), data = nhefs)

## STEP2. Fit Outcome Regression E[Y|A=a, L=l, R]
p.qsmk.obs <- ifelse(nhefs$qsmk == 0, 1-predict(fit, type = "response"),
                     predict(fit, type = "response"))

nhefs$w <- 1/p.qsmk.obs
summary(nhefs$w)  # Min 1.06, Max 16.01 / Mean 2.00 
```

```{r}
# R = W^A if A = 1 ; R = -W^A if A = 0 
nhefs$r <- ifelse(nhefs$qsmk == 1, 1/p.qsmk.obs, -(1/p.qsmk.obs)) 
summary(nhefs$r)  # -2.92 ~ 16.01
```

```{r}
# GEE, Y ~ A+R, weight <- IPW weight 
msm.w <- geeglm(wt82_71 ~ qsmk + r, data=nhefs, id=seqn, weights = w,  
                corstr="independence")
summary(msm.w)
```

```{r}
## STEP3. Use the predicted values from the outcome model to obtain Standardized mean and outcomes 
nhefs$predicted.meanY <- predict(msm.w, nhefs)  # y hat 
summary(nhefs$predicted.meanY[nhefs$cens==0])
```

```{r}
# create a dataset with 3 copies of each subject
nhefs$interv <- -1     # 1st copy: equal to original one

interv0 <- nhefs       # 2nd copy: treatment set to 0, outcome to missing
interv0$interv <- 0    # treatment A = 0 으로 고정 
interv0$qsmk <- 0
interv0$predicted.meanY <- NA

interv1 <- nhefs       # 3rd copy: treatment set to 1, outcome to missing
interv1$interv <- 1    # treatment A = 1 으로 고정 
interv1$qsmk <- 1
interv1$predicted.meanY <- NA

onesample <- rbind(nhefs, interv0, interv1)  # combining datasets : 1629*3 = 4887 
```

```{r}
std <- glm(predicted.meanY ~ qsmk + sex + race ,
           data=onesample) # 3개의 dataset 으로 modeling 

onesample$predicted_meanY <- predict(std, onesample)
mean(onesample[which(onesample$interv==-1),]$predicted_meanY) 

EY.a0.c0<-mean(onesample[which(onesample$interv==0),]$predicted_meanY)
EY.a1.c0<-mean(onesample[which(onesample$interv==1),]$predicted_meanY)
EY.a1.c0 - EY.a0.c0
```


### (4) Treatment FALSE / Outcome FALSE 


```{r}
nhefs <- read_csv("nhefs.csv")

# 82년도에 관측치가 없어 절단된 데이터 : censoring 
nhefs$cens <- ifelse(is.na(nhefs["wt82"]), 1, 0)
```

```{r}
## STEP1. IPW for treatment A 
# Estimation of P(A=1|L) (=IP weights) : via logistic regression 

# A : qsmk (treatment)
fit <- glm(qsmk ~ sex + race + age + I(age^2) +
           as.factor(education) + smokeintensity +
           I(smokeintensity^2) + smokeyrs + I(smokeyrs^2), 
           family = binomial(), data = nhefs)

## STEP2. Fit Outcome Regression E[Y|A=a, L=l, R]
p.qsmk.obs <- ifelse(nhefs$qsmk == 0, 1-predict(fit, type = "response"),
                     predict(fit, type = "response"))

nhefs$w <- 1/p.qsmk.obs
summary(nhefs$w)  # Min 1.06, Max 17.52 / Mean 2.00 
```

```{r}
# R = W^A if A = 1 ; R = -W^A if A = 0 
nhefs$r <- ifelse(nhefs$qsmk == 1, 1/p.qsmk.obs, -(1/p.qsmk.obs)) 
summary(nhefs$r)  # -2.94 ~ 17.52
```

```{r}
# GEE, Y ~ A+R, weight <- IPW weight 
msm.w <- geeglm(wt82_71 ~ qsmk + r, data=nhefs, id=seqn, weights = w,  
                corstr="independence")
summary(msm.w)
```

```{r}
## STEP3. Use the predicted values from the outcome model to obtain Standardized mean and outcomes 
nhefs$predicted.meanY <- predict(msm.w, nhefs)  # y hat 
summary(nhefs$predicted.meanY[nhefs$cens==0])
```

```{r}
# create a dataset with 3 copies of each subject
nhefs$interv <- -1     # 1st copy: equal to original one

interv0 <- nhefs       # 2nd copy: treatment set to 0, outcome to missing
interv0$interv <- 0    # treatment A = 0 으로 고정 
interv0$qsmk <- 0
interv0$predicted.meanY <- NA

interv1 <- nhefs       # 3rd copy: treatment set to 1, outcome to missing
interv1$interv <- 1    # treatment A = 1 으로 고정 
interv1$qsmk <- 1
interv1$predicted.meanY <- NA

onesample <- rbind(nhefs, interv0, interv1)  # combining datasets : 1629*3 = 4887 
```

```{r}
std <- glm(predicted.meanY ~ qsmk + sex + race ,
           data=onesample) # 3개의 dataset 으로 modeling 

onesample$predicted_meanY <- predict(std, onesample)
mean(onesample[which(onesample$interv==-1),]$predicted_meanY) 

EY.a0.c0<-mean(onesample[which(onesample$interv==0),]$predicted_meanY)
EY.a1.c0<-mean(onesample[which(onesample$interv==1),]$predicted_meanY)
EY.a1.c0 - EY.a0.c0
```

[ACE] Treatment, Outcome - TT 3.3 / FT 3.28 / TF 3.15 / FF 3.17  

Treatment 혹은 Outcome 둘 중 하나만 맞은 경우에도 결과를 잘 나타내 주는 것을 확인할 수 있다. 



## 2. IPW 

```{r}
nhefs <- read_csv("nhefs.csv")

# 82년도에 관측치가 없어 절단된 데이터 : censoring 
nhefs$cens <- ifelse(is.na(nhefs["wt82"]), 1, 0)
```

```{r}
# IPW for treatment A 
# Estimation of P(A=1|L) (=IP weights) : via logistic regression 

# A : qsmk (treatment)
fit <- glm(qsmk ~ sex + race + age + I(age^2) +
             as.factor(education) + smokeintensity +
             I(smokeintensity^2) + smokeyrs + I(smokeyrs^2) +
             as.factor(exercise) + as.factor(active) + wt71 + I(wt71^2),
             family = binomial(), data = nhefs)

# qsmk = 0 이면 1-p(확률값) / qsmk=1 이면 p 
p.qsmk.obs <- ifelse(nhefs$qsmk == 0, 1 - predict(fit, type = "response"),
                     predict(fit, type = "response"))

```

```{r}
# IPW 
nhefs$w <- 1/p.qsmk.obs  # 1629 개의 weight : 1/{P(Ai=1|Li)}, 1/{P(Ai=0|Li)}
summary(nhefs$w)  # 1.056 ~ 16 
```

16 값을 가진 사람은 다른 사람들보다 조금 특이하므로, 한 사람의 data 가 16 배로 부풀려져서 pseudo population 에 들어간다.  
이를 통해 두 집단의 비율이 맞춰진다.

```{r}
# GEE (Generalized Estimating Equations) 

# Y ~ A, weight <- IPW weight 
msm.w <- geeglm(wt82_71 ~ qsmk, data=nhefs, weights=w, id=seqn,
                corstr="independence")
summary(msm.w) 
```

E(Y|A=0) = 1.75 / E(Y|A=1) = 1.75+3.52 : difference causal effect 의 estimator 값 

```{r}
beta <- coef(msm.w) ; SE <- coef(summary(msm.w))[,2]
lcl <- beta-qnorm(0.975)*SE ; ucl <- beta+qnorm(0.975)*SE
cbind(beta, lcl, ucl) 
```

나이대를 맞춰 보았더니, 3.5 년 정도로 격차가 더 벌어지게 된다.  

## 3. Standardization 

```{r}
nhefs <- read_csv("nhefs.csv")

# 82년도에 관측치가 없어 절단된 데이터 : censoring 
table(nhefs$qsmk, is.na(nhefs["wt82"]))
nhefs$cens <- ifelse(is.na(nhefs["wt82"]), 1, 0)
```

```{r}
# \Sigma E(Y|L=l, A=a) * P(L=l)
# data : NHEFS from uncensored individuals (C=0)

fit <- glm(wt82_71 ~ qsmk + sex + race + age + I(age*age) + as.factor(education)
           + smokeintensity + I(smokeintensity*smokeintensity) + smokeyrs
           + I(smokeyrs*smokeyrs) + as.factor(exercise) + as.factor(active)
           + wt71 + I(wt71*wt71) + qsmk*smokeintensity, data=nhefs)
           
# summary(fit)$coef  # coef 
```

```{r}
nhefs$predicted.meanY <- predict(fit, nhefs)  # y hat 
summary(nhefs$predicted.meanY[nhefs$cens==0]) # Min -10.88, Max 9.88 / Mean 2.64 
```

```{r}
# create a dataset with 3 copies of each subject
nhefs$interv <- -1 # 1st copy: equal to original one

interv0 <- nhefs   # 2nd copy: treatment set to 0, outcome to missing
interv0$interv <- 0    # treatment A = 0 으로 고정 
interv0$qsmk <- 0
interv0$wt82_71 <- NA

interv1 <- nhefs   # 3rd copy: treatment set to 1, outcome to missing
interv1$interv <- 1    # treatment A = 1 으로 고정 
interv1$qsmk <- 1
interv1$wt82_71 <- NA

onesample <- rbind(nhefs, interv0, interv1) # combining datasets : 1629*3 = 4887 
```

```{r}
std <- glm(wt82_71 ~ qsmk + sex + race + age + I(age*age)
           + as.factor(education) + smokeintensity
           + I(smokeintensity*smokeintensity) + smokeyrs
           + I(smokeyrs*smokeyrs) + as.factor(exercise)
           + as.factor(active) + wt71 + I(wt71*wt71) + I(qsmk*smokeintensity),
           data=onesample) # 3개의 dataset 으로 modeling 
# summary(std) # coef 

onesample$predicted_meanY <- predict(std, onesample)
```

```{r}
mean(onesample[which(onesample$interv==-1),]$predicted_meanY) # 2.56

EY.a0.c0<-mean(onesample[which(onesample$interv==0),]$predicted_meanY)
EY.a1.c0<-mean(onesample[which(onesample$interv==1),]$predicted_meanY)
EY.a1.c0 - EY.a0.c0  # 3.52 
```

